---
title: "STOKE DATA Maturity"
author: "Analysis By Suzan Nakasendwa"
date: "Last modified: 7th Sept 2026"
output:
  word_document: 
    toc: true
    fig_width: 7
    fig_height: 5
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: false
  pdf_document:
    toc: true
editor_options: 
  chunk_output_type: console
---


<style>
h1.title {text-align: center;margin-bottom: 30px;}
h4.author {text-align: center;margin-top: 10px;margin-bottom: 20px;}
h4.date {text-align: center;margin-top: 10px;margin-bottom: 40px;}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      comment = NA,
                      message = FALSE,
                      warning = FALSE)
```


```{r, include=FALSE}

library(readxl)
Data = read_excel("G:/My Drive/PhD/Manuscripts/4. Completeness stroke data/Raw datasets/Data_ecosystem.xlsx")

View(Data)

```

### Create a new dataset that Defines scores

```{r}

library(dplyr)
# Response-score lookup
score_lookup = c(
  "No" = 0,
  "Yes" = 1, 

  "Strongly disagree" = 0,
  "Disagree" = 1,
  "Neutral" = 2,
  "Agree" = 3,
  "Strongly agree" = 4,

  "Not at all" = 0,
  "To a limited extent" = 1,
  "Moderately" = 2,
  "Strongly" = 3,

  "Partially" = 1,
  "Fully" = 2,

  "Never" = 0,
  "Sometimes" = 1,
  "Often" = 2,

  "Occasionally" = 1,
  "Regularly" = 2,

  "Ad-hoc" = 1,
  "Annually" = 2,
  "More than once per year" = 3,

  "Paper-based only" = 0,
  "Paper with partial electronic entry" = 1,
  "Fully electronic at point of care" = 2,
  "Fully electronic and system integrated" = 3,

  "Not interoperable" = 0,
  "Limited manual exchange" = 1,
  "Partial automated exchange" = 2,
  "Fully interoperable" = 3,

  "Fully paper-based" = 0,
  "Hybrid paper-electronic" = 1,
  "Fully electronic but standalone" = 2,
  "Fully electronic and interoperable" = 3,

  "Descriptive statistics only" = 0,
  "Descriptive statistics only Trend or comparative analysis" = 1)

maturity_score = maturity_data %>%
  mutate(across(-Facility,
                ~unname(score_lookup[as.character(.)])))

```

# Sum scores per domian per facility

```{r}


maturity_score = maturity_score %>%
  mutate(`People and culture` = rowSums(select(., `Personnel responsible for stroke data management`:`Accountability for stroke data quality`)),
         `Data activities` = rowSums(select(., `How is stroke data primarily collected?`:`Stroke data used for clinical care, reporting, research, planning`)),
         `Business process` = rowSums(select(., `Stroke data collection & reporting embedded into routine clinical workflows`:`Long-term plan to sustain stroke data systems`)),
         `Technology` = rowSums(select(., `Level of digitization that best describes your stroke data systems`:`Dashboards/automated reports available for monitoring stroke indicators`)),
         OverallScore = `People and culture` +
           `Data activities` + `Business process` +
           Technology)

# Percentage scores
maturity_score = maturity_score %>%
  mutate(`PC%` = (`People and culture`/ 13) * 100,
         `DA%` = (`Data activities`/ 10) * 100,
         `BP%` = (`Business process`/ 9) * 100,
         `TC%` = (`Technology`/ 11) * 100)

maturity_score = maturity_score %>%
  mutate(`Overall%` = (OverallScore / 43) * 100)

# Then weight scores
maturity_score = maturity_score %>%
  mutate(PC_Weight = (`People and culture` / 13) *30,
         DA_Weight = (`Data activities` / 10) * 25,
         BP_Weight = (`Business process` / 9) * 30,
         TC_Weight = (`Technology` / 11) * 15,
         Overall_Weight = (PC_Weight + DA_Weight +
                            BP_Weight + TC_Weight))

```

### Maturity results

```{r}
maturity_results = maturity_score %>%
  select(Facility,
         `People and culture`:`Overall_Weight`) %>%
  pivot_longer(cols = -Facility, names_to = "Variable",
               values_to = "Score") %>%
  pivot_wider(names_from = Facility,
              values_from = Score) %>%
  mutate(across(where(is.numeric), ~ round(.x, 1)))

library(kableExtra)
kable(maturity_results)


```

### Prepare data for plotting

```{r}

### Define the data (You can source data directly from your dataset)

fac = tibble::tribble(
  ~facility, ~ownership,  ~pc,   ~da,  ~bp,    ~tech,
  "A",       "PFP",       69.2,  50.0, 66.7,   63.6,
  "B",       "Public",    69.2,  40.0, 77.8,   81.8,
  "C",       "Public",    76.9,  30.0, 100.0,  36.4,
  "D",       "Public",    76.9,  70.0, 77.8,   54.5,
  "E",       "Public",    61.5,  80.0, 44.4,   72.7,
  "F",       "PNFP",      61.5,  70.0, 66.7,   81.8) %>%
  mutate(sdms     = round((pc + da + bp + tech) / 4, 1),
    weighted = round(pc * .30 + da * .25 + bp * .30 + 
                       tech * .15, 1),
    ownership = factor(ownership,
                       levels = c("Public", "PNFP", "PFP")))

dom_levels = c("People & culture", "Data activities",
                "Business processes", "Technology")
dom_weight = c(`People & culture` = .30,
                `Data activities` = .25,
                `Business processes` = .30,
                `Technology` = .15)

fac_long = fac %>%
  pivot_longer(c(pc, da, bp, tech), names_to = "dom",
               values_to = "score") %>%
  mutate(domain = factor(recode(dom,
      pc = "People & culture", da = "Data activities",
      bp = "Business processes", tech = "Technology"),
      levels = dom_levels))

# Define Shared theme, palette and maturity bands 
own_cols = c(Public = "#2C6E9B", PNFP = "#E1812C",
             PFP = "#3A923A")
bands = tibble::tibble(lower = c(0, 12.5, 37.5, 62.5, 87.5),
                       upper = c(12.5, 37.5, 62.5, 87.5, 100),
                       label = c("Chaotic", "Reactive", "Stable",
                                 "Proactive", "Predictive"),
                       fill  = c("#F5F5F5", "#EDEDED", "#F5F5F5",
                                 "#EDEDED", "#F5F5F5"))

library(ggplot2)
theme_mat = theme_minimal(base_size = 10) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(colour = "#444444", fill = NA,
                                    linewidth = .4),
        plot.title = element_text(size = 11, hjust = 0),
        plot.caption = element_text(size = 7.4,hjust = 0,
                                    colour = "#555555"),
    legend.position  = "top",
    legend.title     = element_blank(),
    legend.key.size  = unit(.4, "cm"))


```

# Figure 3 : Domain maturity by ownership 

```{r}

grp = bind_rows(fac_long %>% group_by(ownership, domain) %>%
                  summarise(mean = mean(score), .groups = "drop"),
                fac %>% group_by(ownership) %>%
                  summarise(mean = mean(sdms),
                            .groups = "drop") %>%
                  mutate(domain = "Overall (SDMS)")) %>%
  mutate(domain = factor(domain, levels = c(dom_levels,
                                            "Overall (SDMS)")))

pts = bind_rows(fac_long %>% select(ownership, domain, score),
                fac %>% transmute(ownership, domain = "Overall (SDMS)", score = sdms)) %>%
  filter(ownership == "Public") %>%
  mutate(domain = factor(domain, levels = c(dom_levels,
                                            "Overall (SDMS)")))

n_lab = fac %>% count(ownership) %>%
  mutate(lab = paste0(ownership, " (n=", n, ")")) %>% 
  pull(lab, ownership)

fig3 = ggplot() +
  geom_rect(data = bands, aes(xmin = -Inf, xmax = Inf,
                              ymin = lower, ymax = upper),
            fill = bands$fill) +
  geom_hline(yintercept = c(12.5, 37.5, 62.5, 87.5),
             linetype = "dashed",
             colour = "#9A9A9A", linewidth = .3) +
  geom_col(data = grp, aes(domain, mean, fill = ownership),
           position = position_dodge(width = .75), width = .70,
           colour = "white", linewidth = .3) +
  geom_point(data = pts, aes(x = as.numeric(domain) - 0.26,
                             y = score),
             position = position_dodge(width = .78),
             size = 1.1,
             colour = "#222222", alpha = .75) +
  geom_text(data = grp, aes(domain, mean, group = ownership,
                            label = sprintf("%.1f", mean)),
            position = position_dodge(width = .78),
            vjust = -.6, size = 2.5) +
   annotate("text", x = 5.47,
            y = (bands$lower + bands$upper)/2,
           label = bands$label, angle = 90, 
           hjust = 0.5, size = 2.5,
           colour = "#8A8A8A", fontface = "bold.italic") +
  scale_fill_manual(values = own_cols, labels = n_lab) +
  scale_y_continuous(limits = c(0, 100),
  breaks = c(0, 12.5, 37.5, 62.5, 87.5, 100),
  expand = expansion(mult = c(0, 0.02))) +
  labs(x = NULL, y = "Standardised maturity score (%)",
       caption = paste("PNFP - Private not-for-profit; PFP - Private for-profit; PNFP and PFP each comprise a single facility.\nBars are group means; black points are individual public facilities (n=4).")) + theme_mat +
  theme(axis.title.y = element_text(colour = "black"),
        axis.text.x  = element_text(colour = "black",
                                    face = "bold"),
        axis.text.y  = element_text(colour = "black",
                                    face = "bold"))

fig3

```

### Figure 4 : unweighted vs weighted, against maturity bands

```{r}

dumb = fac %>% 
  mutate(ownership = factor(
    ownership,levels = c("Public", "PNFP", "PFP")),
    facility_lab = paste0("Facility ", facility,
                          "  (", ownership, ")")) %>%
  arrange(ownership, sdms) %>%
  mutate(facility_lab = factor(facility_lab, levels = rev(facility_lab)))

fig4 = ggplot(dumb) +
  geom_rect(data = bands, aes(ymin = -Inf, ymax = Inf,
                              xmin = lower, xmax = upper),
            fill = bands$fill, inherit.aes = FALSE) +
  geom_vline(xintercept = c(12.5, 37.5, 62.5, 87.5),
             linetype = "dashed",
             colour = "#9A9A9A", linewidth = .3) +
  geom_segment(aes(x = sdms, xend = weighted,
                   y = facility_lab, yend = facility_lab),
               colour = "#BBBBBB", linewidth = 1) +
  geom_point(aes(sdms, facility_lab, colour = ownership,
                 shape = "Unweighted SDMS"),
             size = 3, fill = "white",
             stroke = 1.1) +
  geom_point(aes(weighted, facility_lab, colour = ownership,
                 fill = ownership,
                 shape = "Weighted score"),
             size = 3, stroke = 1.1) +
  geom_text(aes(sdms, facility_lab,
                label = sprintf("%.1f", sdms)),
            vjust = -1.1, size = 2.5, colour = "#444444") +
  geom_text(aes(weighted, facility_lab,
                label = sprintf("%.1f", weighted)),
            vjust = 1.7, size = 2.5, colour = "#444444",
            fontface = "bold") +
  annotate("text", x = (bands$lower + bands$upper) / 2,
           y = 6.5,label = bands$label, size = 2.5,
           colour = "#8A8A8A", fontface = "bold.italic") +
  scale_colour_manual(values = own_cols) +
  scale_fill_manual(values = own_cols) +
  scale_x_continuous(limits = c(0, 100),
  breaks = c(0, 12.5, 37.5, 62.5, 87.5, 100),
  expand = expansion(mult = c(0, 0.02))) +
  labs(x = "Maturity score (%)", y="",
       caption = paste("PNFP - Private not-for-profit, PFP - Private for-profit\nWeighting as per domain weights: People & culture - 30%, Data activities - 25%, Business processes - 30%,\nTechnology - 15%),")) +
  theme_mat +
  theme(plot.margin = margin(10,8, 8, 5),
        axis.title.y = element_text(colour = "black"),
        axis.title.x = element_text(colour = "black"),
        axis.text.x  = element_text(colour = "black",
                                    face = "bold"),
        axis.text.y  = element_text(colour = "black",
                                    face = "bold")) +
  scale_shape_manual(values = c("Unweighted SDMS" = 21,
                                "Weighted score" = 21),
                     guide = guide_legend(override.aes = list(
                       fill = c("white", "black"),
                       colour = "black")))

fig4

```

### Figure 5 : which domain drives the weighted score

```{r}

drivers = fac_long %>%
  group_by(ownership, domain) %>%
  summarise(mean = mean(score), .groups = "drop") %>%
  mutate(available = dom_weight[as.character(domain)],
         points    = mean * available,
         seg_lab   = sprintf("%.1f / %d", points, round(available * 100)))

totals = drivers %>% group_by(ownership) %>%
  summarise(total = sum(points), .groups = "drop")

fig5 = ggplot(drivers, aes(ownership, points, fill = domain)) +
  geom_col(width = .55, colour = "white", linewidth = .4) +
  geom_text(aes(label = seg_lab),
            position = position_stack(vjust = .5),
            colour = "white", fontface = "bold", size = 2.6) +
  geom_text(data = totals,aes(ownership, total, 
                              label = sprintf("%.1f%%",
                                              total)),
            inherit.aes = FALSE, vjust = -.7, 
            fontface = "bold", size = 3.2) +
  scale_fill_manual(values = c(
    "People & culture"   = "#4C72B0",
    "Data activities"    = "#DD8452",
    "Business processes" = "#55A868",
    "Technology"         = "#8172B3")) +
  scale_x_discrete(labels = n_lab) +
  scale_y_continuous(limits = c(0, 80),expand = c(0, 0)) +
  labs(x = NULL, y = "Weighted score\n (percentage points)", caption = paste("PNFP - Private not-for-profit; PFP - Private for-profit")) +
  theme_mat +
  theme(axis.title.y = element_text(colour = "black"),
        axis.title.x = element_text(colour = "black"),
        axis.text.x  = element_text(colour = "black",
                                    face = "bold"),
        axis.text.y  = element_text(colour = "black",
                                    face = "bold"))

fig5

```

# Figure 2: Item-level matrix

```{r}

items = tibble::tribble(
  ~item,  ~domain,  ~max, ~A, ~B, ~C, ~D, ~E, ~F, ~invariant,
  "F1  Designated stroke-data personnel",   "People & culture",   1, 1,1,0,0,0,1, FALSE,
  "F2  Staff see stroke data as critical",  "People & culture",   4, 0,3,4,4,4,2, FALSE,
  "F3  Routine staff training",             "People & culture",   3, 3,1,1,1,0,2, FALSE,
  "F4  Data used for audit / QI",           "People & culture",   1, 1,1,1,1,1,0, FALSE,
  "F5  Leadership promotes data use",       "People & culture",   3, 3,2,3,3,2,2, FALSE,
  "F6  Accountability for data quality",    "People & culture",   1, 1,1,1,1,1,1, TRUE,
  "F7  Mode of data capture",               "Data activities",    3, 1,1,0,3,3,2, FALSE,
  "F8  Standardised definitions / ICD",     "Data activities",    1, 1,1,1,1,1,1, TRUE,
  "F9  Routine data-quality checks",        "Data activities",    2, 2,1,1,1,2,2, FALSE,
  "F10 Level of analysis performed",        "Data activities",    3, 0,0,0,1,1,1, FALSE,
  "F11 Data reused for >1 purpose",         "Data activities",    1, 1,1,1,1,1,1, TRUE,
  "F12 Embedded in clinical workflow",      "Business processes", 2, 2,2,2,1,1,1, FALSE,
  "F13 SOPs for stroke data",               "Business processes", 1, 0,1,1,0,0,1, FALSE,
  "F14 Data reviewed in QI meetings",       "Business processes", 2, 2,1,2,2,1,1, FALSE,
  "F15 Formal governance structures",       "Business processes", 1, 0,1,1,1,1,1, FALSE,
  "F16 Data inform planning / policy",      "Business processes", 2, 1,1,2,2,1,2, FALSE,
  "F17 Long-term sustainability plan",      "Business processes", 1, 1,1,1,1,0,0, FALSE,
  "F18 Level of digitisation",              "Technology",         3, 1,1,0,1,1,1, FALSE,
  "F19 Interoperability with other systems","Technology",         3, 2,3,1,2,2,3, FALSE,
  "F20 Unique patient identifier",          "Technology",         1, 0,1,1,0,1,1, FALSE,
  "F21 Secure storage",                     "Technology",         1, 1,1,1,1,1,1, TRUE,
  "F22 Secure backup",                      "Technology",         1, 1,1,0,0,1,1, FALSE,
  "F23 Access controls",                    "Technology",         1, 1,1,1,1,1,1, TRUE,
  "F24 Dashboards / automated reports",     "Technology",         1, 1,1,0,1,1,1, FALSE) %>%
  mutate(domain = factor(domain, levels = dom_levels))

# Facility order grouped by ownership
fac_order = c("B", "C", "D", "E", "F", "A")

items_long = items %>%
  pivot_longer(A:F, names_to = "facility", values_to = "score") %>%
  mutate(prop = score / max,
         facility = factor(facility, levels = fac_order),
         item = factor(item, levels = rev(items$item)),
         cell = paste0(score, "/", max))

library(ggh4x)
fig2 = ggplot(items_long, aes(facility, item, fill = prop)) +
  geom_tile(colour = "white", linewidth = .4) +
  geom_text(aes(label = cell), colour = "black",size = 2.3,
            fontface = "bold") +
  geom_vline(xintercept = c(4.5, 5.5), colour = "#333333",
             linewidth = 0.8) +
  facet_grid2(domain ~ .,scales = "free_y", space = "free_y",
              strip = strip_themed(
                text_y = elem_list_text(colour = c(
      "#4C72B0", "#DD8452", "#55A868", "#8172B3"),
      face = c("bold","bold", "bold", "bold"),
      vjust=c(-2,-2,-2,-2)))) +
  scale_fill_gradientn(colours = c(
    "#C0392B", "#E8A33D", "#F5E17A", "#9BCB6B", "#3F8F3F"),
    limits = c(0, 1), 
    name = "Proportion of maximum item score",
    guides(fill = guide_colourbar(
      barheight = unit(6, "cm"), barwidth = unit(0.4, "cm"),
    title.position = "right"))) +
  scale_x_discrete(labels = paste0("Facility ", fac_order),
                   position = "bottom") +
  coord_cartesian(clip = "off") +
  labs(x = NULL, y = NULL, caption = paste("Facilities are grouped by ownership: Public (B, C, D, E) | PNFP (F) | PFP (A).\nPNFP - Private not-for-profit; PFP - Private for-profit")) +
  theme_minimal(base_size = 9) +
  theme(panel.grid = element_blank(),
        strip.placement = "outside",
        strip.text.y.left = element_text(
          angle = 0, face = "bold", size = 8),
        axis.text.y = element_blank(),
        legend.position = "right",
        plot.caption = element_text(
          size = 7.2, colour = "#555555", hjust = 0)) +
  geom_text(data = items_long %>% distinct(item, domain),
            aes(x = 0.4, y = item, label = item, 
                colour = domain),hjust = 1, size = 2.6,
            inherit.aes = FALSE) +
scale_colour_manual(values = c(
  "People & culture" = "#4C72B0",
  "Data activities" = "#DD8452",
  "Business processes" = "#55A868",
  "Technology" = "#8172B3"), guide = "none") +
  theme(plot.margin = margin(5.5, 5.5, 5.5, 130),
        axis.text.y = element_blank(),
        axis.text.x = element_text(colour = "black",
                                   face = "bold"),
        panel.grid = element_blank(),
        legend.position = "right",
  legend.direction = "vertical",
  legend.title = element_text(
    angle = 0,
    hjust = 0
  ))

library(cowplot)
library(grid)

p_no_title = fig2 +
  scale_fill_gradientn(
    colours = c("#C0392B", "#E8A33D", "#F5E17A",
                "#9BCB6B", "#3F8F3F"),
    limits = c(0, 1), name = NULL) +
  guides(fill = guide_colourbar(barheight = unit(7, "cm"),
                       barwidth  = unit(0.35, "cm")))

leg = cowplot::get_legend(
  p_no_title + theme(legend.position = "right",
      legend.margin = margin(-15,0, 0, 10)))

leg_title = grid::textGrob(
  "Proportion of maximum item score", rot = 90,
  vjust=-2, hjust=0.5,
  gp = grid::gpar(fontsize = 9))

custom_legend = cowplot::plot_grid(leg, leg_title, nrow = 1,
                ncol = 2, rel_widths = c(1, 2), align = "v")

fig2_final = cowplot::plot_grid(
  p_no_title + theme(legend.position = "none"),
  custom_legend, nrow = 1, rel_widths = c(1, 0.15),
  align = "h")

fig2_final


```



# Supplemental figure : sensitivity to non-discriminating items

```{r}

sens = items %>%
  pivot_longer(A:F, names_to = "facility", values_to = "score") %>%
  group_by(facility, domain) %>%
  summarise(full = sum(score) / sum(max) * 100,
            trim = sum(score[!invariant]) /
              sum(max[!invariant]) * 100,
            .groups = "drop") %>%
  group_by(facility) %>%
  summarise(reported = round(mean(full), 1),
            trimmed  = round(mean(trim), 1), .groups = "drop") %>%
  # use the published SDMS on the x-axis so the figure matches Table 2
  left_join(fac %>% select(facility, sdms), by = "facility") %>%
  mutate(reported = sdms,
         drops    = trimmed < 62.5 & reported >= 62.5,
         maintained = trimmed < 62.5 & reported < 62.5,
         facility_lab = factor(paste0("Facility ", facility),
                               levels = paste0("Facility ", facility[order(sdms)])))

figS = ggplot(sens) +
  geom_rect(data = bands, aes(ymin = -Inf, ymax = Inf,
                              xmin = lower, xmax = upper),
            fill = bands$fill, inherit.aes = FALSE) +
  geom_vline(xintercept = c(12.5, 37.5, 62.5, 87.5), linetype = "dashed",
             colour = "#9A9A9A", linewidth = .3) +
  geom_segment(aes(x = reported, xend = trimmed,
                   y = facility_lab, yend = facility_lab,
                   colour = ifelse(maintained, "maintained",
                                   as.character(drops))),
               arrow = arrow(length = unit(.16, "cm"),
                             type = "closed"), linewidth = .8) +
  geom_point(aes(reported, facility_lab), shape = 21,
             fill = "white", colour = "#333333",
             size = 2.8, stroke = 1) +
  geom_point(aes(trimmed, facility_lab,
                 colour = ifelse(maintained, "maintained",
                                 as.character(drops))),
             size = 2.8) +
  geom_text(aes(reported, facility_lab,
                label = sprintf("%.1f", reported)),
            hjust = -.35, size = 2.4, colour = "#444444") +
  geom_text(aes(trimmed, facility_lab,
                label = sprintf("%.1f", trimmed),
                colour = ifelse(maintained, "maintained",
                                as.character(drops))),
            hjust = 1.35, size = 2.4, show.legend = FALSE) +
  annotate("text", x = (bands$lower + bands$upper) / 2,
           y = 6.6, label = bands$label,
           size = 2.5, colour = "#8A8A8A",
           fontface = "bold.italic") +
  scale_colour_manual(
  values = c(`FALSE` = "#888888",`TRUE` = "#c08f2f",
             `maintained` = "#c02f2f"),
  labels = c(`FALSE` = "No change in level",
             `TRUE` = "Reclassified to Level 3 (stable)",
             `maintained` = "Maintained")) +
  scale_x_continuous(limits = c(0, 100),
  breaks = c(0, 12.5, 37.5, 62.5, 87.5, 100),
  expand = expansion(mult = c(0, 0))) +
  coord_cartesian(clip = "off", ylim = c(0.5, 6.5)) +
  labs(x = "Stroke Data Maturity Score (%)", y = NULL,
       caption = paste("Hollow point = as reported (24 items); filled point = excluding the five items scored at maximum by all facilities (F6, F8, F11, F21, F23).\nFacilities A and C move from proactive (Level 4) to stable (Level 3).")) +
  theme_mat + theme(plot.margin = margin(24, 8, 8, 8)) +
  theme(axis.title.y = element_text(colour = "black"),
        axis.title.x = element_text(colour = "black"),
        axis.text.x  = element_text(colour = "black",
                                    face = "bold"),
        axis.text.y  = element_text(colour = "black",
                                    face = "bold"))

figS

```

### Supplemental Table

```{r}

kable(Table6_data)

```


